The Galactic Delay-Time Distribution of Type Ia Supernovae:\\ A Chemical Evolution Perspective

### 2022-08-15 src/tex/ms.tex
[showyourwork] overleaf sync

\usepackage{showyourwork}
\usepackage{amsmath}
% \usepackage{siunitx}

@@

\subsection{Accuracy of astroNN}
\label{subsec:astronn-quality}

@@

\subsection{Measurements of the Type Ia Supernova Delay Time Distribution}
\label{subsec:dtd-measurements}

@@

\subsection{One-Zone Models}
\label{subsec:onezone-results}

@@

Adjusting the parameters of the DTD affects how steeply the abundance track in [Fe/H]-[O/Fe] space falls from its initial ``plateau'' and the location of the ``knee.'' Figure \ref{fig:onezone-slope-timescale} shows abundance tracks from one-zone models which assume different slopes of a power-law DTD or different timescales of an exponential DTD. A steeper power-law slope results in a steeper decline in [O/Fe] and a lower knee, as does a shorter exponential timescale. The consequence is that a very shallow power-law produces similar observable signatures to a very steep exponential, which suggests these models may be difficult to distinguish in this parameter space.

\begin{figure}
    \centering
    \includegraphics{figures/onezone_slope_timescale.pdf}
    \caption{\textit{Center:} Abundance tracks in the [Fe/H]-[O/Fe] plane for one-zone models with various DTD parameters. Tracks from models which assume a power-law DTD are in black, with dotted, dashed, and solid curves representing power-law slopes ($\alpha$) of $-0.8$, $-1.1$, and $-1.4$, respectively. Tracks from models which assume an exponential DTD are in blue, with dotted, dashed, and solid curves representing exponential decay timescales ($\tau$) of 1.5 Gyr, 3 Gyr, and 6 Gyr, respectively. In all cases a minimum delay time of 40 Myr is assumed. \textit{Top:} MDFs generated from by one-zone models. \textit{Right:} Distribution of [O/Fe] generated by the one-zone models.}
    \label{fig:onezone-slope-timescale}
    \script{onezone_slope_timescale.py}
\end{figure}

@@

\noindent We use a non-linear least squares method to fit Equation \ref{eq:greggio05-approx} to the analytical DTDs for DD progenitors. We obtain best-fit parameters $\alpha_1=-0.19$, $A_{\rm{rise}}=1.7$, $\tau_{\rm{rise}}=0.084$ Gyr, $\alpha_2=-0.91$, $A_{\rm{tail}}=0.85$, and $\tau_{\rm{tail}}=0.088$ Gyr for the ``WIDE'' scheme, and $\alpha_1=-0.53$, $A_{\rm{rise}}=3.0$, $\tau_{\rm{rise}}=0.042$ Gyr, $\alpha_2=-1.1$, $A_{\rm{tail}}=0.62$, and $\tau_{\rm{tail}}=0.11$ Gyr for the ``CLOSE'' scheme. These best-fit approximations are represented by the dashed curves in Figure \ref{fig:greggio05-dtd}. The difference between the true analytical DTDs and our best-fit approximations is small enough as to be irrelevant for our purposes, so for the rest of this section we rely on the best-fit approximations for faster simulations.

\begin{figure}
    \centering
    \includegraphics{figures/greggio05_dtd.pdf}
    \caption{Analytical DTDs computed by \citet{Greggio2005-AnalyticalRates}.}
    \label{fig:greggio05-dtd}
    \script{greggio05_dtd.py}
\end{figure}

Figure \ref{fig:onezone-greggio05} shows the results of one-zone chemical evolution models with the \citet{Greggio2005-AnalyticalRates} DTDs. We assume $\eta=2.5$, $\tau_*=2$ Gyr, an inside-out SFH evaluated at a radius of 8 kpc, a continuous recycling approximation, and a minimum SN Ia delay of 40 Myr. We compare the SD, DD WIDE, and DD CLOSE schemes to standard power-law and exponential DTDs. The SD and DD CLOSE DTDs follow nearly identical tracks in [O/Fe] vs [Fe/H] to an exponential DTD with a 1.5 Gyr timescale. The DD WIDE DTD likewise follows a similar track to a 3 Gyr exponential.

\begin{figure}
    \centering
    \includegraphics{figures/onezone_greggio05.pdf}
    \caption{Caption}
    \label{fig:onezone-greggio05}
    \script{onezone_greggio05.py}
\end{figure}

>>>>

\usepackage{showyourwork}
\usepackage{amsmath}
\usepackage{layouts}
% \usepackage{siunitx}

@@

\subsection{Accuracy of astroNN}
\label{sec:astronn-quality}

@@

\subsection{Measurements of the Type Ia Supernova Delay Time Distribution}
\label{sec:dtd-measurements}

@@

\subsection{One-Zone Models}
\label{sec:onezone-results}

All models are run for a duration of 13.2 Gyr with a simulation timestep of 0.01 Gyr.

@@

Adjusting the parameters of the DTD affects how steeply the abundance track in [Fe/H]-[O/Fe] space falls from its initial ``plateau'' and the location of the ``knee.'' Figure \ref{fig:onezone-slope-timescale} shows abundance tracks from one-zone models which assume different slopes of a power-law DTD or different timescales of an exponential DTD. A steeper power-law slope results in a steeper decline in [O/Fe] and a lower knee, as does a shorter exponential timescale. The consequence is that a very shallow power-law produces a similar abundance track to a very steep exponential, which suggests these models may be difficult to distinguish in this parameter space. However, these cases can be distinguished by their MDFs and [O/Fe] distributions, where the power-law DTD produces a narrower, less skewed distribution while the exponential DTD produces a broader, more skewed distribution. In the [O/Fe] distribution especially, an exponential DTD with a 6 Gyr timescale produces a large number of high-$\alpha$ stars.

\begin{figure}
    \centering
    \includegraphics{figures/onezone_slope_timescale.pdf}
    \caption{\textit{Center:} Abundance tracks in the [Fe/H]-[O/Fe] plane for one-zone models with various DTD parameters. Tracks from models which assume a power-law DTD are in black, with dotted, dashed, and solid curves representing power-law slopes ($\alpha$) of $-0.8$, $-1.1$, and $-1.4$, respectively. Tracks from models which assume an exponential DTD are in blue, with dotted, dashed, and solid curves representing exponential decay timescales ($\tau$) of 1.5 Gyr, 3 Gyr, and 6 Gyr, respectively. In all cases a minimum delay time of 40 Myr is assumed. \textit{Top:} MDFs generated by the one-zone models. \textit{Right:} Distribution of [O/Fe] generated by the one-zone models.}
    \label{fig:onezone-slope-timescale}
    \script{onezone_slope_timescale.py}
\end{figure}

@@

\noindent We use a non-linear least squares method to fit Equation \ref{eq:greggio05-approx} to the analytical DTDs for DD progenitors. We obtain best-fit parameters $\alpha_1=-0.19$, $A_{\rm{rise}}=1.7$, $\tau_{\rm{rise}}=0.084$ Gyr, $\alpha_2=-0.91$, $A_{\rm{tail}}=0.85$, and $\tau_{\rm{tail}}=0.088$ Gyr for the ``WIDE'' scheme, and $\alpha_1=-0.53$, $A_{\rm{rise}}=3.0$, $\tau_{\rm{rise}}=0.042$ Gyr, $\alpha_2=-1.1$, $A_{\rm{tail}}=0.62$, and $\tau_{\rm{tail}}=0.11$ Gyr for the ``CLOSE'' scheme.\footnote{
Our implementation of these best-fit approximations, as well as the original SD and DD formulations, is available in the GitHub repository for this paper.
} These best-fit approximations are represented by the dashed curves in Figure \ref{fig:greggio05-dtd} \citep[see also Figure 8 in][]{Greggio2005-AnalyticalRates}. The difference between the true analytical DTDs and our best-fit approximations is small enough as to be irrelevant for our purposes, so for the rest of this section we rely on the best-fit approximations for faster simulations.

\begin{figure}
    \centering
    \includegraphics{figures/greggio05_dtd.pdf}
    \caption{Analytical DTDs computed by \citet{Greggio2005-AnalyticalRates}. The solid red curve represents the SD DTD for Chandrasekhar explosions, while the green and blue solid curves represent, respectively, the CLOSE and WIDE formulations of a DD DTD. The best-fit approximations for the DD DTDs (see Equation \ref{eq:greggio05-approx}) are represented by the dashed olive and cyan curves, respectively. All DTDs are normalized to unity.}
    \label{fig:greggio05-dtd}
    \script{greggio05_dtd.py}
\end{figure}

Figure \ref{fig:onezone-greggio05} shows the results of one-zone chemical evolution models with the \citet{Greggio2005-AnalyticalRates} DTDs. We assume $\eta=2.5$, $\tau_*=2$ Gyr, an inside-out SFH evaluated at a radius of 8 kpc, a continuous recycling approximation, and a minimum SN Ia delay of 40 Myr. We compare the SD, DD WIDE, and DD CLOSE schemes to standard power-law, broken power-law, and exponential DTDs. The SD and DD CLOSE DTDs follow nearly identical tracks in [O/Fe] vs [Fe/H]; however, their distributions on [O/Fe] differ at the low end. The SD DTD follows an exponential with a 1.5 Gyr timescale, whereas the DD CLOSE DTD is well-approximated by a broken power-law with an initial plateau of 350 Myr and a subsequent declining slope of -1.1. The WIDE prescription is likewise best approximated by a broken power-law, but with a longer plateau width of 1 Gyr. In all cases, the difference between the analytical DTD and the nearest broken power-law or exponential is likely too small to be observationally detectable, so in our multi-zone models we implement the more generic functions in lieu of the analytical DTDs.

A multi-zone with the \citet{Greggio2005-AnalyticalRates} SD DTD produced nearly identical results to an exponential DTD with a 1.5 Gyr timescale.

\begin{figure*}
    \centering
    \includegraphics{figures/onezone_greggio05_single.pdf}
    \includegraphics{figures/onezone_greggio05_double.pdf}
    \caption{Abundance tracks and distributions from one-zone models with the analytical DTDs from \citet{Greggio2005-AnalyticalRates}. \textit{Left:} a comparison between the analytical SD DTD (solid curve), an exponential DTD with a 1.5 Gyr timescale (dashed curve), and a power-law DTD with a slope of -1.1 (dotted curve). \textit{Right:} a comparison between two different prescriptions for an analytical DD DTD (solid curves), power-law DTDs with an initial plateau of 1 Gyr and 350 Myr (dashed curves), and a power-law DTD with no plateau and a slope of -1.1 (dotted curve).}
    \label{fig:onezone-greggio05}
    \script{onezone_greggio05_single.py}
    \script{onezone_greggio05_double.py}
\end{figure*}

\subsection{Multi-Zone Models}
\label{sec:multizone-results}
